---
title: '"Visualisation avec ggplot"'
author: "Aissa Morin"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE)
```

# Introduction

Dans cet exercice nous allons utiliser le package `ggplot2` pour visualiser des données 

# Données

On va utiliser les données de la base de données PanTHERIA

# Exercice

```{r}
library(tidyverse)
```

##Lecture des données

```{r, echo = FALSE, eval = FALSE}
#pantheria <- readr::read_delim("../data/pantheria-traits/PanTHERIA_1-0_WR05_Aug2008.txt", delim = "\t")
# Ou
# readr::read_delim("https://raw.githubusercontent.com/FRBCesab/datatoolbox/master/data/PanTHERIA_1-0_WR05_Aug2008.txt", delim = "\t")

# pantheria <- readr::read_delim(here::here("data","pantheria-traits", "PanTHERIA_1-0_WR05_Aug2008.txt"), delim = "\t")
```

__Nettoyage des données avec une fonction__

```{r}
pantheria  <- data_pantheria()# fichier data_pantheria.R
```

## Tidy the dataset 

Nettoyage des données
```{r, echo = FALSE, eval = FALSE}

# dat <- pantheria %>%
  # mutate(order = forcats::as_factor(MSW05_Order),
  #        family = forcats::as_factor(MSW05_Family)) %>%
  # rename( adult_bodymass  = '5-1_AdultBodyMass_g',
  #         dispersal_age  = '7-1_DispersalAge_d',
  #         gestation = '9-1_GestationLen_d',
  #         homerange = '22-2_HomeRange_Indiv_km2',
  #         litter_size = '16-1_LittersPerYear',
  #         longevity = '17-1_MaxLongevity_m') %>%
  # select(order, family, adult_bodymass, dispersal_age, gestation,homerange,litter_size,longevity ) %>%
  # na_if(-999 )
  # 
  # dat
  
```

__Nettoyage des données avec une fonction__

```{r}
dat  <- clean_pantheria(pantheria) #fichier wriggle.R
```


## Data Exploration 

How many observation in family ? in order ?
```{r}

dat %>% 
  count(order)

dat %>% 
  count(family)
  
dat %>%
  filter(!is.na(homerange))%>%
  group_by(family)%>%
  summarise(m = mean(homerange), sd =  sd(homerange, n = n()))

```

## Graph 1: Bar chart of the number of entries per family

```{r}

dat %>%
  group_by(family) %>% 
  mutate(n = n()) %>% 
  filter(n > 100) %>%
  
  ggplot()+
  geom_col()+
  aes(forcats::fct_reorder(family, n), x = n) +
  xlab("Family") +
  ylab("Counts") +
  ggtitle("Number of entries per family")

  

```

## Graph 2: Scatter plot of litter size as a function of longevity

```{r}

theme_set(theme_bw()) # play around with theme
dat %>%
  filter(!is.na(litter_size), !is.na(longevity)) %>% #Filter out missing values in both litter size and longevity
  group_by(family) %>% # group by family
  mutate(n = n()) %>% # count the number of entries per family
  mutate(longevity = longevity / 12) %>% # Change month to year
  filter(n > 10) %>% # select only those families with more than 10 entries
  ggplot() +
  aes(x = longevity, y = litter_size, col = family) + # scatter plot
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) + # se = TRUE to add confidence intervals
  xlab("Longevity") + # add label for X axis
  ylab("Litter size") + # add label for Y axis
  ggtitle("Scatterplot") + # add title
  facet_wrap(~ family, nrow = 3) # split in several panels,
                                 # one for each family
                                 # remove scale = 'free' for
                                 # same scale for all plots

```




